Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-10-29 paper/main.tex
Added short text on mutable vs immutable. Needs more work.
Added figure of an UEB and some text on capabilities in relation to hashing.

\usepackage[numbers]{natbib}
\usepackage{appendix} % Fixes formatting of appendices
\usepackage[printonlyused]{acronym} % Package to handle the acronym list
\usepackage{graphicx} % We *may* use images
\graphicspath{{images/}} % and it is clean to put them in a separate dir
\usepackage{hyperref} % Internal and external links is nice
\hypersetup{pdfborder=0 0 0} % ..especially without red borders

@@

\begin{acronym}
\acro{AES}{Advanced Encryption Standard}
\acro{FEC}{Forward Error Correction}
\acro{LAFS}{Least-Authority Filesystem}
\acro{NIST}{National Institute of Standards and Technology}
\acro{RAM}{Random Access Memory}
\acro{SHA}{Secure Hash Algorithm}
\acro{SSE2}{Streaming SIMD Extensions 2}
\acro{SSSE3}{Supplemental Streaming SIMD Extensions 3}
\acro{SUPERCOP}{System for Unified Performance Evaluation Related to
Cryptographic Operations and Primitives}
\end{acronym}

@@

There are two kinds of files in the Tahoe-\ac{LAFS} -- \textbf{immutable} and
\textbf{mutable} files.

@@

The properties of erasure coding can be thought of as those of parity bits in
RAID\footnote{Redundant Array of Independent Disks} systems. \citet*{t_erasure}
compare erasure coding and plain replication, and conclude that ``\emph{...
erasure codes have mean time to failures many orders of magnitude higher than
replicated systems with similar storage and bandwidth requirements.}''

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=0.9\columnwidth]{Tahoe-hashing.pdf}
    \caption{Tahoe-LAFS: Location of hashing.}
    \label{fig:tahoehashing}
    \emph{Figure from
     \href{http://tahoe-lafs.org/source/tahoe/trunk/docs/specifications/mut.svg}
     {tahoe-lafs.org}}
\end{figure}

As seen in figure \ref{fig:tahoehashing}, the usage of secure hashing is
extensive in Tahoe, and has to be considered as a key part of the functionality.

@@

A hash tree, also known as \emph{Merkle tree}, is a type of data structure
which can be described as a tree with nodes that can verify all information
below in the hierarchy, as depicted in figure \ref{fig:tahoemerkletree}. This
enables Tahoe to verify small segments of a file at the time, and this can be
used for instance to start playing a movie file while it is still being
downloaded.

\paragraph{Encryption key.}

\paragraph{Ciphertext and shares.}

Stored in Capability/URI Extension Block.
URI Extension Block $=$ Root hash $+$ file size $+$ encoding parameters.
URI / File read cap = Hash of UEB $+ K_c$

@@

file-encoding1.svg

>>>>

\usepackage[numbers]{natbib} % We need to tweak our referencing a bit.
\usepackage{appendix} % Fixes formatting of appendices
\usepackage[printonlyused]{acronym} % Package to handle the acronym list
\usepackage{graphicx} % We *may* use images
\graphicspath{{images/}} % and it is clean to put them in a separate dir
\usepackage{hyperref} % Internal and external links is nice
\hypersetup{pdfborder=0 0 0} % ..especially without red borders

@@

\begin{acronym}
\acro{AES}{Advanced Encryption Standard}
\acro{CEB}{Capability Extension Block}
\acro{FEC}{Forward Error Correction}
\acro{LAFS}{Least-Authority Filesystem}
\acro{NIST}{National Institute of Standards and Technology}
\acro{RAM}{Random Access Memory}
\acro{RAID}{Redundant Array of Independent Disks}
\acro{SHA}{Secure Hash Algorithm}
\acro{SSE2}{Streaming SIMD Extensions 2}
\acro{SSSE3}{Supplemental Streaming SIMD Extensions 3}
\acro{SUPERCOP}{System for Unified Performance Evaluation Related to
Cryptographic Operations and Primitives}
\acro{UEB}{URI Extension Block}
\end{acronym}

@@

There are two kinds of files in the Tahoe-\ac{LAFS} -- \textbf{immutable} and
\textbf{mutable} files. An immutable file is created exactly once, i.e. it
cannot be modified, and can be read repeatedly. Mutable files can be modified,
and everyone who has access to the \emph{signing key} can make new versions of
the mutable file.
%More text here?

@@

The properties of erasure coding can be thought of as those of replication in
\acsu{RAID} systems. \citet*{t_erasure} compare erasure coding and plain
replication, and conclude that ``\emph{...  erasure codes have mean time to
failures many orders of magnitude higher than replicated systems with similar
storage and bandwidth requirements.}''

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.9\columnwidth]{Tahoe-hashing.pdf}
    \caption{Tahoe-LAFS: Hashing operations in mutable files.}
    \label{fig:tahoehashing}
    \emph{Figure from
     \href{http://tahoe-lafs.org/source/tahoe/trunk/docs/specifications/mut.svg}
     {tahoe-lafs.org}}
\end{figure}

As seen in figure \ref{fig:tahoehashing}, the usage of secure hashing is
extensive in Tahoe, and has to be considered as a key part of the functionality,
and thus affecting the performance.

@@

A hash tree, also known as a \emph{Merkle tree}, is a type of data structure
which can be described as a tree with nodes that can verify all information
below in the hierarchy, as depicted in figure \ref{fig:tahoemerkletree}. This
enables Tahoe to verify small segments of a file at the time, and this can be
used for instance to start playing a movie file while it is still being
downloaded.

These secure hashes of the shares of a file, are contained within a small
datastructure named the \emph{\ac{CEB}}.

\paragraph{Capabilities.}

A \textbf{capability} (or an URI) contains the encryption key, and a hash of the
\emph{\ac{UEB}}. The \ac{UEB} for each file is a data structure containing the
hash of the \ac{CEB}, the size of the file, any encoding parameters necessary to
perform the erasure decoding, and a hash of the plaintext and the encrypted
text. This is illustrated in figure \ref{fig:tahoeueb}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.9\columnwidth]{Tahoe-UEB.pdf}
    \caption{The composition of an \ac{UEB}.}
    \label{fig:tahoeueb}
\end{figure}

@@

A hash of the encryption key is used to form the "storage index", which is
used for both server selection and to index shares within each storage node.
% More here?
